---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
master.data.all <- read.csv("activity.csv", stringsAsFactors=FALSE, header=TRUE)
master.data <- na.omit(master.data.all)
```

## Histogram of total number of steps taken each day
```{r}
hist(master.data$steps,col="green",main="Total Number of steps taken each day")
```



## What is mean total number of steps taken per day?
```{r}
library(plyr)
mean.steps <- ddply(master.data, .(date), summarize, mean_value=mean(steps))
hist(mean.steps$mean_value,col="green",main="Mean Total steps taken per day")
abline(v=median(mean.steps$mean_value),col="magenta",lwd=2)
```


## What is the average daily activity pattern?
```{r}
mean.interval <- ddply(master.data, .(interval), summarize, mean_value=mean(steps))
with(mean.interval, plot(interval,mean_value,main="Average Daily Activity Pattern",type="l"))
```

```{r computetime,echo=FALSE}
maxsteps <- mean.interval$interval[which(mean.interval$mean_value==max(mean.interval$mean_value))]
```

The 5 minute interval on average across all the days in the dataset that contains the maximum number of steps is `r maxsteps`

## Imputing missing values
```{r computevals,echo=FALSE}
missingvalues <- sum(is.na(master.data.all$steps))
```

Total number of missing values in the dataset is `r missingvalues`


```{r}
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
master.data.all <- ddply(master.data.all, ~ interval, transform, steps = impute.mean(steps))

mean.steps <- ddply(master.data.all, .(date), summarize, mean_value=mean(steps))
hist(mean.steps$mean_value,col="green")
abline(v=median(mean.steps$mean_value),col="magenta",lwd=2)

```



## Are there differences in activity patterns between weekdays and weekends?

```{r}
master.data.all$dayframe <- ifelse(weekdays(as.Date(master.data.all$date))%in%c("Saturday","Sunday"),"Weekend","Weekday")
mean.dayframe <- ddply(master.data.all, .(interval), summarize, mean_value=mean(steps))
with(mean.dayframe, plot(interval,mean_value,main="Average Daily Activity Pattern",type="l"))
```


